---
title: "STAT 480 Project"
author: "Kailee Ervin"
date: "4/6/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(knitr)
library(rmarkdown)
library(ggplot2)
library(forcats)
library(readr)
library(dplyr)
library(tidyr)
library(lubridate)
library(stringr)
library(scales)
library(zoo)
library(tidyverse)

my_theme <- function (base_size = 12, base_family = "Times New Roman") {
  theme_gray() +
    theme(
      rect = element_rect(fill = NA, linetype = 0, colour = NA, size = base_size/22),
      text = element_text(family = base_family, face = "plain", color = "black", size = base_size, hjust = 0.5, vjust = 0.5, angle = 0, lineheight = 0.9, margin = margin(), debug = FALSE),
      panel.border = element_blank(),
      panel.background = element_blank(),
      plot.caption = element_text(color = "grey50"),
      legend.position = "right",
      panel.grid.minor.y = element_line(colour = "grey80"),
      panel.grid.major.x = element_line(colour = "grey80", size = rel(0.25)),
      panel.grid.minor = element_line(size = rel(0.25), colour = "grey80", linetype = "dotted"),
      panel.grid.major.y = element_line(colour = "grey80", size = rel(0.3)),
      plot.title = element_text(face = "bold", size = base_size * 1.5)
    )
}
theme_set(my_theme())
my_pal <- c("#16A08CFF", "plum3", "gold", "#EE0011FF", "#149BEDFF", "lightpink", "lightsalmon", "#A1C720FF", "#FA6B09FF", "paleturquoise", "seagreen", "palevioletred4", "lightpink", "navajowhite2", "lightblue3", "deepskyblue", "coral", "aquamarine")
scale_colour_discrete <- function(...) {scale_colour_manual(..., values = my_pal)}
scale_fill_discrete <- function(...) {scale_fill_manual(..., values = my_pal)}
```

Download Data
```{r}
youtube<-read_csv("https://raw.githubusercontent.com/frost2351/stat480/master/youtube.csv?token=AO7D46QTNVIF4KNJJZT2RUK6UDEE2")
```


# Data cleaning
```{r}
youtube<-youtube %>% 
  mutate(video_id=as.character(video_id),
         trending_date=ydm(as.character(trending_date)),
         title=as.character(title),
         channel_title=as.character(channel_title),
         category_id=recode(as.factor(category_id), "1"="Film & Animation", "2"="Autos & Vehicles", "10"="Music",
                            "15"="Pets and Animals", "17"="Sports", "19"="Travel & Events", "20"="Gaming",
                            "22"="People and Blogs", "23"="Comedy", "24"="Entertainment", 
                            "25"="News and Politics", "26"="Howto & Style", "27"="Education", 
                            "28"="Science & Technology", "29"="Nonprofits & Activism", "30"="Movies",
                            "43"="Shows", "44"="Trailers"),
         publish_date=dmy(as.character(publish_date)),
         time_frame=factor(time_frame, levels=c("0:00 to 0:59","1:00 to 1:59","2:00 to 2:59","3:00 to 3:59",
                                                "4:00 to 4:59","5:00 to 5:59","6:00 to 6:59","7:00 to 7:59",
                                                "8:00 to 8:59","9:00 to 9:59","10:00 to 10:59","11:00 to 11:59",
                                                "12:00 to 12:59","13:00 to 13:59","14:00 to 14:59",
                                                "15:00 to 15:59","16:00 to 16:59","17:00 to 17:59",
                                                "18:00 to 18:59","19:00 to 19:59","20:00 to 20:59",
                                                "21:00 to 21:59","22:00 to 22:59","23:00 to 23:59")),
         published_day_of_week=factor(published_day_of_week, levels=c("Monday", "Tuesday", "Wednesday",
                                                                      "Thursday", "Friday", "Saturday",
                                                                      "Sunday")),
         tags=as.character(tags))
```

# Number of Trending Videos By Publish Country
```{r}
youtube %>% 
  group_by(publish_country) %>% 
  mutate(total_video=n()) %>% 
  distinct(video_id, .keep_all=TRUE) %>% 
  summarise(`Distinct Videos`=n(),
            `Total Obs`=mean(total_video)) %>% 
  pivot_longer(cols=`Distinct Videos`:`Total Obs`,
               names_to="total_type",
               values_to="number_videos") %>% 
  ggplot(aes(x=total_type, fill=publish_country))+
  scale_fill_discrete(name="Publish Country")+
  geom_bar(aes(weight=number_videos), show.legend = FALSE)+
  facet_wrap(~publish_country)+
  labs(title="Total Number of Videos \nvs. \nTotal Distinct Videos By Country", y="Number of Videos", x="Type of Total")
```


# Are there differences in publishing times throughout the day for trending videos?
```{r}
# youtube_time<-youtube %>% 
#   mutate(hour=recode(time_frame, "0:00 to 0:59"="1","1:00 to 1:59"="2","2:00 to 2:59"="3",
#                                           "3:00 to 3:59"="4","4:00 to 4:59"="5","5:00 to 5:59"="6",
#                                           "6:00 to 6:59"="7","7:00 to 7:59"="8","8:00 to 8:59"="9",
#                                           "9:00 to 9:59"="10","10:00 to 10:59"="11","11:00 to 11:59"="12",
#                                           "12:00 to 12:59"="13","13:00 to 13:59"="14","14:00 to 14:59"="15",
#                                           "15:00 to 15:59"="16","16:00 to 16:59"="17","17:00 to 17:59"="18",
#                                           "18:00 to 18:59"="19","19:00 to 19:59"="20","20:00 to 20:59"="21",
#                                           "21:00 to 21:59"="22","22:00 to 22:59"="23","23:00 to 23:59"="24"),
#          hour=as.character(hour),
#          hour=as.integer(hour)) %>% 
#   group_by(publish_country) %>% 
#   distinct(video_id, .keep_all=TRUE) %>% 
#   mutate(distinct_videos=n()) %>% 
#   group_by(publish_country, hour) %>% 
#   summarise(time_frame_total=n(),
#             distinct_videos=mean(distinct_videos)) %>% 
#   mutate(time_frame_prop=case_when(publish_country=="CANADA"~time_frame_total/distinct_videos,
#                                       publish_country=="FRANCE"~time_frame_total/distinct_videos,
#                                       publish_country=="GB"~time_frame_total/distinct_videos,
#                                       publish_country=="US"~time_frame_total/distinct_videos))
# MEAN PUBLISH HOUR:
# CANADA	14.81067: 15			
# FRANCE	15.77379		: 16	
# GB	13.93609: 14	
# US	14.76669: 15			
youtube %>% 
  mutate(hour=recode(time_frame, "0:00 to 0:59"="1","1:00 to 1:59"="2","2:00 to 2:59"="3",
                                          "3:00 to 3:59"="4","4:00 to 4:59"="5","5:00 to 5:59"="6",
                                          "6:00 to 6:59"="7","7:00 to 7:59"="8","8:00 to 8:59"="9",
                                          "9:00 to 9:59"="10","10:00 to 10:59"="11","11:00 to 11:59"="12",
                                          "12:00 to 12:59"="13","13:00 to 13:59"="14","14:00 to 14:59"="15",
                                          "15:00 to 15:59"="16","16:00 to 16:59"="17","17:00 to 17:59"="18",
                                          "18:00 to 18:59"="19","19:00 to 19:59"="20","20:00 to 20:59"="21",
                                          "21:00 to 21:59"="22","22:00 to 22:59"="23","23:00 to 23:59"="24"),
         hour=as.character(hour),
         hour=as.integer(hour)) %>% 
  group_by(publish_country) %>% summarise(mean_hour=mean(hour, na.rm=TRUE))


# boxplot
youtube %>% 
  mutate(hour=recode(time_frame, "0:00 to 0:59"="1","1:00 to 1:59"="2","2:00 to 2:59"="3",
                                          "3:00 to 3:59"="4","4:00 to 4:59"="5","5:00 to 5:59"="6",
                                          "6:00 to 6:59"="7","7:00 to 7:59"="8","8:00 to 8:59"="9",
                                          "9:00 to 9:59"="10","10:00 to 10:59"="11","11:00 to 11:59"="12",
                                          "12:00 to 12:59"="13","13:00 to 13:59"="14","14:00 to 14:59"="15",
                                          "15:00 to 15:59"="16","16:00 to 16:59"="17","17:00 to 17:59"="18",
                                          "18:00 to 18:59"="19","19:00 to 19:59"="20","20:00 to 20:59"="21",
                                          "21:00 to 21:59"="22","22:00 to 22:59"="23","23:00 to 23:59"="24"),
         hour=as.character(hour),
         hour=as.integer(hour)) %>%
  ggplot(aes(x=publish_country, y=hour))+ 
  geom_boxplot(aes(fill=publish_country), show.legend=FALSE)+
  scale_y_continuous(breaks=seq(0,24,2))+
  scale_x_discrete(limits=c("US", "GB", "FRANCE", "CANADA"))+
  coord_flip()+
  labs(title="Trending Videos Publishing Hour \nby Country", x="Publish Country", y="Hour")
?ylim
# # Distribution of Publish Time by Country
# ggplot(youtube_time, aes(x=hour, fill=publish_country))+
#   scale_fill_discrete(name="Publish Country")+
#   geom_bar(aes(weight=time_frame_prop), show.legend = FALSE)+
#   facet_grid(publish_country~.)+
#  labs(x="Hour", y="Proportion of Total Distinct Videos \nPublished By Country", title="Distribution of Times \nWhen Trending Videos are Published")
  
```

```{r}
youtube %>% 
  mutate(hour=recode(time_frame, "0:00 to 0:59"="1","1:00 to 1:59"="2","2:00 to 2:59"="3",
                                          "3:00 to 3:59"="4","4:00 to 4:59"="5","5:00 to 5:59"="6",
                                          "6:00 to 6:59"="7","7:00 to 7:59"="8","8:00 to 8:59"="9",
                                          "9:00 to 9:59"="10","10:00 to 10:59"="11","11:00 to 11:59"="12",
                                          "12:00 to 12:59"="13","13:00 to 13:59"="14","14:00 to 14:59"="15",
                                          "15:00 to 15:59"="16","16:00 to 16:59"="17","17:00 to 17:59"="18",
                                          "18:00 to 18:59"="19","19:00 to 19:59"="20","20:00 to 20:59"="21",
                                          "21:00 to 21:59"="22","22:00 to 22:59"="23","23:00 to 23:59"="24"),
         hour=as.character(hour),
         hour=as.integer(hour)) %>% 
  ggplot(aes(x=published_day_of_week, y=hour))+ 
  geom_boxplot(aes(fill=publish_country), show.legend=FALSE)+scale_y_continuous(breaks=seq(0,24,2))+facet_grid(publish_country~.)+coord_flip()+
  labs(title="Trending Videos Publishing Hour Each Day of Week \nby Country", x="Day of Week", y="Hour")



youtube %>% 
  mutate(hour=recode(time_frame, "0:00 to 0:59"="1","1:00 to 1:59"="2","2:00 to 2:59"="3",
                                          "3:00 to 3:59"="4","4:00 to 4:59"="5","5:00 to 5:59"="6",
                                          "6:00 to 6:59"="7","7:00 to 7:59"="8","8:00 to 8:59"="9",
                                          "9:00 to 9:59"="10","10:00 to 10:59"="11","11:00 to 11:59"="12",
                                          "12:00 to 12:59"="13","13:00 to 13:59"="14","14:00 to 14:59"="15",
                                          "15:00 to 15:59"="16","16:00 to 16:59"="17","17:00 to 17:59"="18",
                                          "18:00 to 18:59"="19","19:00 to 19:59"="20","20:00 to 20:59"="21",
                                          "21:00 to 21:59"="22","22:00 to 22:59"="23","23:00 to 23:59"="24"),
         hour=as.character(hour),
         hour=as.integer(hour)) %>% 
  ggplot(aes(x=published_day_of_week, y=hour))+ 
  geom_boxplot(aes(fill=published_day_of_week))+scale_y_continuous(breaks=seq(0,24,2))+coord_flip()+
  labs(title="Trending Videos Publishing Hour Each Day of Week", x="Day of Week", y="Hour")
  
```



# Publishing Time throughout week by country?
```{r}
# By Week by Country
youtube %>% 
  group_by(publish_country) %>% 
  distinct(video_id, .keep_all=TRUE) %>% 
  mutate(distinct_videos=n()) %>% 
  group_by(publish_country, published_day_of_week) %>% 
  summarise(time_frame_total=n(),
            distinct_videos=mean(distinct_videos)) %>% 
  mutate(time_frame_prop=case_when(publish_country=="CANADA"~time_frame_total/distinct_videos,
                                      publish_country=="FRANCE"~time_frame_total/distinct_videos,
                                      publish_country=="GB"~time_frame_total/distinct_videos,
                                      publish_country=="US"~time_frame_total/distinct_videos)) %>% 
  ggplot(aes(x=published_day_of_week, fill=publish_country))+
  scale_fill_discrete(name="Publish Country")+
  geom_bar(aes(weight=time_frame_prop), show.legend=FALSE)+
  facet_grid(publish_country~.)+
  labs(x="Day of Week", y="Proportion of Total Distinct Videos Published By Country", title="Days of Week When Trending Videos Published", subtitle="by Country")+
  coord_flip()+
  theme(text = element_text(family = "Times New Roman", color = "grey20"))

```


# Length Between Publish Date and Trending Date
```{r}
youtube_trend<-youtube %>% 
  select(video_id, publish_date, trending_date) %>% 
  group_by(video_id, publish_date) %>% 
  summarise(first_trend=min(trending_date)) %>% 
  mutate(days_to_trend=(publish_date %--% first_trend)/ddays(1)) %>% 
  ungroup()
# mean days to trend:7.129273	
youtube_trend %>% summarise(mean_days_to_trend=mean(days_to_trend))

# No Trendline
youtube_trend %>% 
  ggplot(aes(x=publish_date, y=days_to_trend, color=days_to_trend))+
  scale_color_continuous(name ="Days to Trending")+
  geom_point(aes(alpha=0.4))+
  labs(title="Days from Published to Trending", x="Publish Date", y="Days to Trending")

#Trendline
youtube_trend %>% 
  ggplot(aes(x=publish_date, y=days_to_trend, color=days_to_trend))+
  scale_color_continuous(name ="Days to Trending")+
  geom_smooth(method = "lm", color="grey20", alpha=0.3, size=0.7, se=FALSE)+
  geom_point(aes(alpha=0.4))+
  labs(title="Days from Published to Trending", x="Publish Date", y="Days to Trending")


```

```{r}
#adjusting linear line
youtube_old<-youtube_trend %>% 
  filter(publish_date<"2017-11-14")

youtube_trend %>% 
  filter(publish_date<"2017-11-14") %>% 
  summarise(n=n())
# videos before november 14, 2017: 1115	
youtube_trend %>% 
  summarise(n=n())
# total videos: 56106	

  ggplot(data=youtube_old, aes(x=publish_date, y=days_to_trend, color=days_to_trend))+
  scale_color_continuous(name ="Days to Trending")+
  geom_smooth(data=youtube_old, method = "lm", color="grey20", alpha=0.3, size=0.7, se=FALSE)+
  geom_point(aes(alpha=0.4))+
  labs(title="Days from Published to Trending", subtitle="July 23, 2006 to November 14, 2017",x="Publish Date", y="Days to Trending")
```


```{r}
#Looking at publish dates before november 14
youtube_old %>% 
  ggplot(aes(x=publish_date, y=days_to_trend, color=days_to_trend))+
  scale_color_continuous(name ="Days to Trending")+
  geom_smooth(method = "lm", color="grey20", alpha=0.3, size=0.7, se=FALSE)+
  geom_point(aes(alpha=0.4))+
  labs(title="Days from Published to Trending", x="Publish Date", y="Days to Trending")
```


```{r}
youtube_trend %>% 
  filter(publish_date>="2017-11-14") %>% 
  ggplot(aes(x=publish_date, y=days_to_trend, color=days_to_trend))+
  scale_color_continuous(name="Days to Trending")+
  geom_point(aes(alpha=0.4))+
  labs(title="Days from Published to Trending",subtitle="from November 14, 2017 to June 14, 2018", x="Publish Date", y="Days to Trending")

youtube_trend %>% 
  filter(publish_date>="2017-11-14") %>% summarise(mean_days_to_trend=mean(days_to_trend))

```


# What is the weird blank line in the middle????
```{r}
youtube %>% 
  arrange(trending_date) %>% group_by(trending_date) %>% summarise(nbr_trending_videos=n()) %>% filter(trending_date>="2018-04-01", trending_date<="2018-04-30") %>% slice(1:10)
```

# Average Number of trending videos in each country each day (For Intro)
```{r}
youtube %>% group_by(publish_country, trending_date) %>% summarise(n=n()) %>% group_by(publish_country) %>% summarise(mean_videos=mean(n))
```
Publish Date
2006-07-23	
2018-06-14

Trending Date
2017-11-14	
2018-06-14	
```{r}
youtube %>% group_by(trending_date) %>% summarise(n=n()) %>% arrange(trending_date)
```

