---
title: "Exploring the Trends of Trending Videos: A Youtube Analysis"
author: "Taylor Buske, Kailee Ervin, Derek Thada, Saiyothin Vongpanya"
date: "4/22/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(knitr)
library(rmarkdown)
library(ggplot2)
library(forcats)
library(readr)
library(dplyr)
library(tidyr)
library(lubridate)
library(stringr)
library(extrafont)
library(maps)
library(reactable)

my_theme <- function (base_size = 12, base_family = "Times New Roman") {
  theme_gray() +
    theme(
      rect = element_rect(fill = NA, linetype = 0, colour = NA, size = base_size/22),
      text = element_text(family = base_family, face = "plain", color = "black", size = base_size, hjust = 0.5, vjust = 0.5, angle = 0, lineheight = 0.9, margin = margin(), debug = FALSE),
      panel.border = element_blank(),
      panel.background = element_blank(),
      plot.caption = element_text(color = "grey50"),
      legend.position = "right",
      panel.grid.minor.y = element_line(colour = "grey80"), 
      panel.grid.major.x = element_line(colour = "grey80", size = rel(0.25)),
      panel.grid.minor = element_line(size = rel(0.25), colour = "grey80", linetype = "dotted"),
      panel.grid.major.y = element_line(colour = "grey80", size = rel(0.3)),
      plot.title = element_text(face = "bold", size = base_size * 1.5)
    )
}
theme_set(my_theme())
my_pal <- c("#16A08CFF", "plum3", "gold", "#EE0011FF", "#149BEDFF", "lightpink", "lightsalmon", "#A1C720FF", "#FA6B09FF", "paleturquoise", "seagreen", "palevioletred4", "lightpink", "navajowhite2", "lightblue3", "deepskyblue", "coral", "aquamarine")
scale_colour_discrete <- function(...) {scale_colour_manual(..., values = my_pal)}
scale_fill_discrete <- function(...) {scale_fill_manual(..., values = my_pal)}
```

## Introduction
About our data:
```{r, include=F}
youtube <- read_csv("https://raw.githubusercontent.com/frost2351/stat480/master/youtube.csv")
str(youtube)
```
Our YouTube data set has 161,470 records and 17 variables. The variables in this dataset are: "video_id", "trending_date", "title, "channel_title", "category_id"  "publish_date", "time_frame", "published_day_of_week", "publish_country", "tags", "views", "likes", "dislikes", "comment_count", "comments_disabled", "ratings_disabled", and "video_error_or_removed". The variables have various data types that we will be using in our analysis, including character, integer, boolean, and factors. Each row in the data set is a particular trending video on a specific trending date. Additionally, four separate countries are analyzed, the United States, Canada, Great Britain, and France. Each country has its own list of trending videos on each day. The trending videos are taken from November of 2017 to June of 2018. According to Google (the owner of YouTube), the trending list is updated approximately every 15 minutes **(Citation Needed)**. Thus, the number of videos that are trending throughout a day fluctuates. The number of videos on the trending list at any given time is around 200 in each country.


## Questions
Our questions focus on five main areas of focus:

* Video Categories
* Date and Times
* Titles
* Duration and Metrics
* Locations

## Data Cleaning
For the most part, our data set is quite user friendly. When loading the data, R automatically assigns certain data types. However, some of the automatic data types assigned are not helpful for future analysis and were changed. 

* Trending_date: changed from character to date
* Publish_date: changed from character to date
* Time_frame: factor reordered so that first level is "0:00 to 0:59" and last level is "23:00 to 23:59"
* Published_day_of_week: factor reordered so that week starts of “Monday” and ends on “Sunday”

We also had to clean the category_id variable. The raw data assigns a number to each category. We researched what categories these numbers corresponded to and relabeled the data using the category names. Using Youtube's API (<https://gist.github.com/dgp/1b24bf2961521bd75d6c>), we relabeled the numbers to factors.


## Categories:

```{r, echo=FALSE}
youtube <- youtube %>% 
  mutate(video_id = as.character(video_id),
         trending_date = ydm(as.character(trending_date)),
         title = as.character(title),
         channel_title = as.character(channel_title),
         category_id = recode(as.factor(category_id), "1"="Film & Animation", "2"="Autos & Vehicles",
                              "10"="Music", "15"="Pets and Animals", "17"="Sports", 
                              "19"="Travel &Events", "20"="Gaming", "22"="People and Blogs", 
                              "23"="Comedy", "24"="Entertainment", 
                              "25"="News and Politics", "26"="Howto & Style", "27"="Education",
                              "28"="Science & Technology", "29"="Nonprofits & Activism", "30"="Movies",
                              "43"="Shows", "44"="Trailers"),
         publish_date = dmy(as.character(publish_date)),
         time_frame=factor(time_frame, levels=c("0:00 to 0:59","1:00 to 1:59","2:00 to 2:59","3:00 to 3:59",
                                                "4:00 to 4:59","5:00 to 5:59","6:00 to 6:59","7:00 to 7:59",
                                                "8:00 to 8:59","9:00 to 9:59","10:00 to 10:59","11:00 to 11:59",
                                                "12:00 to 12:59","13:00 to 13:59","14:00 to 14:59",
                                                "15:00 to 15:59","16:00 to 16:59","17:00 to 17:59",
                                                "18:00 to 18:59","19:00 to 19:59","20:00 to 20:59",
                                                "21:00 to 21:59","22:00 to 22:59","23:00 to 23:59")),
         published_day_of_week=factor(published_day_of_week, levels=c("Monday", "Tuesday", "Wednesday",
                                                                      "Thursday", "Friday", "Saturday",
                                                                      "Sunday")),
         tags = as.character(tags))

youtube$time_frame1 <- fct_collapse(youtube$time_frame, morning = c("4:00 to 4:59","5:00 to 5:59",
                                                                    "6:00 to 6:59","7:00 to 7:59",
                                                                    "8:00 to 8:59","9:00 to 9:59",
                                                                    "10:00 to 10:59","11:00 to 11:59"), 
                                    afternoon = c("12:00 to 12:59","13:00 to 13:59","14:00 to 14:59",
                                                  "15:00 to 15:59","16:00 to 16:59", "17:00 to 17:59",
                                                  "18:00 to 18:59","19:00 to 19:59"), 
                                    night = c("20:00 to 20:59","21:00 to 21:59","22:00 to 22:59",
                                              "23:00 to 23:59", "0:00 to 0:59","1:00 to 1:59",
                                              "2:00 to 2:59","3:00 to 3:59"))

youtube$published_day_of_week1 <- fct_collapse(youtube$published_day_of_week, 
                                               weekday = c("Monday", "Tuesday", "Wednesday", "Thursday"),
                                               weekend = c("Friday", "Saturday", "Sunday"))
```

### Do different categories have differing instances on being on the trending page?

Our data consists of 18 different video categories. These categories are broad and range from “Pets and Animals” to “Music” to “News and Politics”. Figure 1 shows the number of trending videos in each category in our data.

```{r Figure 1}
youtube %>% 
  group_by(category_id) %>% 
  mutate(n = n()) %>% 
  ggplot(aes(x = fct_reorder(category_id, n),
             fill = category_id)) + 
  geom_bar(show.legend = FALSE) +
  labs(x = "Category", 
       y = "Instances on Trending Page", 
       title = "Instances on Trending Page in each Category",
       caption = "Figure 1") +
  coord_flip() 
```

Thus, we can see that “Entertainment” videos are the most frequent type of videos that appear on the trending list. It is important to note that each video can only be listed under one category.


We can also see how the number of trending videos in each category changes over time as well. Our data consists of trending videos from November of 2017 through June of 2018. Figure 2 below shows the number of trending videos over time in the categories “Entertainment,” “Music,” “People and Blogs,” “Comedy,” and “News and Politics”. These categories are the top five categories with the most trending videos as shown above. We aim to answer the following question: 

How has the number of trending videos for different categories change over time?
```{r Figure 2}
left_right <- youtube %>% 
  filter(category_id %in% c("Entertainment", "Music",
                            "People and Blogs", "Comedy",
                            "News and Politics")) %>%
  group_by(category_id, trending_date) %>%
  tally() %>% 
  group_by(category_id) %>% 
  slice(which.min(trending_date), which.max(trending_date))

youtube %>% 
  filter(category_id %in% c("Entertainment", "Music",
                            "People and Blogs", "Comedy",
                            "News and Politics")) %>%
  group_by(category_id, trending_date) %>%
  tally() %>% 
  ggplot() +
  geom_line(aes(x = trending_date,
                y = n,
                color = fct_reorder(category_id, desc(n)))) +
  geom_point(data = left_right,
             aes(x = trending_date,
                 y = n,
                 color = category_id)) +
  geom_line(data = left_right,
            aes(x = trending_date,
                y = n,
                color = category_id),
            linetype = "dashed",
            size = 0.40) +
  labs(x = "Date", 
       y = "Instances on Trending Page", 
       title = "Instances on Trending Page in each Category Time Series",
       color = "Category",
       caption = "Observing the top five categories \n Figure 2") +
  scale_fill_discrete(name = "Category") 

```

In analyzing Figure 2, we see that the number of trending videos for “Entertainment” stays relatively constant over time. For “Music” videos, we see an increase starting around March and this increase continues into May. The increase in “Music” makes sense as many artists release music during this time (so the song can become popular before summertime, but is still considered new). The categories of “Comedy,” “People and Blogs,” and “News and Politics” are constant throughout until “Music” begins to make its increase. During “Music”’s increase, these three categories decrease.


One metric that YouTube uses to determine which videos are trending is the amount of views the video has. Figure 3 below shows the top five categories with the most views on their videos.

### How do people view trending videos for different categories?

```{r Figure 3}
youtube %>%
  group_by(video_id) %>% 
  slice(which.max(trending_date)) %>% 
  group_by(category_id) %>%
  summarise(total_views = sum(views, na.rm = T)) %>%
  arrange(desc(total_views)) %>%
  top_n(5) %>%
  ggplot(aes(x = fct_reorder(category_id, desc(total_views)),
             fill = category_id)) +
  geom_bar(aes(weight=total_views),
           show.legend = FALSE) +
  labs(x = "Categories", 
       y = "Total Views", 
       title = "Top 5 Categories by Views",
       caption = "Considering the views from the last day trending \n Figure 3") +
  scale_y_continuous(labels = scales::comma)
```
In analyzing Figure 3, we see that “Music” has over double the amount of views as the next highest category, “Entertainment.” Earlier, we found that the most frequent type of trending video was “Entertainment,” so this result seems unintuitive. We would expect to see “Entertainment” with the most views since this category has the most trending videos. However, according to YouTube, new music releases are some of the most popular videos on the site, in terms of views. Additionally, popular music takes time to produce and is not released as frequently as entertainment videos, which is an extremely broad category. So, while the quantity of “Entertainment” videos exceeds “Music”, the popularity (by views) for “Music” exceeds “Entertainment.” Another explanation for why “Music” has more views even though “Entertainment” has more videos is because we think people are more likely to replay “Music” videos. People can listen to music over and over again without staring at their screen. “Entertainment” videos generally require people to actually take the time to view them. In looking at the most viewed videos in YouTube history, 18 out of the top 20 are “Music” videos.


## How do people comment on trending videos for different categories?
```{r Figure 4}
youtube %>%
  group_by(video_id) %>% 
  slice(which.max(trending_date)) %>% 
  group_by(category_id) %>%
  summarise(total_comments = sum(comment_count)) %>%
  arrange(desc(total_comments)) %>%
  top_n(5) %>%
  ggplot(aes(x = fct_reorder(category_id, desc(total_comments)),
             fill = category_id)) +
  geom_bar(aes(weight = total_comments),
           show.legend = FALSE) +
  labs(x = "Categories", 
       y = "Total Comments", 
       title = "Top 5 Categories by Comments",
       caption = "Considering the comments from the last day trending \n Figure 4") +
  scale_y_continuous(labels = scales::comma)
```

```{r Comments per View}
youtube %>%
  group_by(category_id) %>%
  summarise(total_comments = sum(comment_count),
            total_views = sum(views, na.rm = T),
            rate = total_comments/total_views) %>%
  filter(category_id %in% c("Music", "Entertainment")) %>%
  arrange(desc(rate)) %>%
  select(-total_comments,-total_views)

```

## First Impressions: Titles and Their Effects on Trending Videos

The first thing we dived into was looking 
### Do different countries have different title length in their trending videos?

```{r Figure 5}

youtube %>% 
  distinct(video_id, .keep_all = TRUE) %>% 
  mutate(title_length = str_length(title)) %>% 
  ggplot(aes(x = title_length, 
             y = ..ndensity.., 
             fill = publish_country)) +
  geom_histogram(binwidth=10,
                 show.legend = F) +
  facet_wrap(~publish_country) +
  labs(x = "Length of Title", 
       y = "Proportion of Trending Videos", 
       title = "Length of Titles in Trending Youtube Videos \nin Different Countries",
       caption = "Calculated with distinct video ID's") 

```
This is necessary because we need to see if we need

### Do different countries use differing amount of capital letters in their trending videos?

One thing we had to be careful about is correctly scaling our data to better represent we wanted to visualize. Since the length of videos in each country can be different Since the United States and Great Britain have a significantly lower number of distinct videos, we used proportion of 
```{r Figure 5}
youtube %>% 
  distinct(video_id, .keep_all = TRUE) %>% 
  mutate(avg_capitals_per_title_length = str_count(title, "[A-Z]") / str_length(title)) %>% 
  ggplot(aes(x = avg_capitals_per_title_length, 
             y = ..ndensity.., 
             fill = publish_country)) +
  geom_histogram(binwidth=0.03,
                 show.legend = F) +
  facet_wrap(~publish_country) +
  labs(x = "Proportion of Capital Letters in Title", 
       y = "Proportion of Trending Videos", 
       title = "Proportion of Capital Letters in Titles of Trending \nYoutube Videos in Different Countries",
       caption = "Calculated with distinct video ID's")

youtube %>% 
  distinct(video_id, .keep_all = TRUE) %>% 
  mutate(avg_capitals_per_title_length = str_count(title, "[A-Z]") / str_length(title)) %>% 
  ggplot(aes(x = avg_capitals_per_title_length,
             y = views)) +
  geom_point(aes(col=publish_country)) + 
  geom_smooth(method="loess", se=F)


```
From Figure 5, we can see that Canada, Great Britain, and the United States have similar peaks. France is a clear outlier in its distribution. 


```{r}
youtube %>% 
  distinct(video_id, .keep_all = TRUE) %>% 
  mutate(words = str_extract_all(title, "[A-z]{1,}"),
         wordcount = lengths(words)) %>% 
  ggplot(aes(x = wordcount, 
             y = ..ndensity.., 
             fill = publish_country)) +
  geom_histogram(binwidth=1,
                 show.legend = F) +
  facet_wrap(~publish_country) +
  labs(x = "Number of Words", 
       y = "Proportion of Trending Videos", 
       title = "Number of Words in Titles of Trending Youtube Videos\n in Different Countries",
       caption = "Calculated with distinct video ID's")
  


```
### Have trending videos seen a change in their usage of capital letters?
```{r more Capital Letters}
youtube %>% 
  distinct(video_id, .keep_all = TRUE) %>% 
  mutate(number_capitals = str_count(title, "[A-Z]"),
         year = year(trending_date)) %>% 
  ggplot(aes(x = number_capitals,
             y = ..ndensity..)) +
  geom_histogram(binwidth = 1, 
                 fill = "plum3",
                 color = "black",
                 size = 0.1) +
  facet_wrap(~year) +
  labs(x = "Number of Capital Letters in Title",
       y = "Proportion of Trending Videos",
       title = "Number of Capital Letters in Trending Video Titles \nbetween 2017 and 2018",
       caption = "Calculated with distinct video ID's")

```

### Is there a discernible trend in the usage of exclamation marks?
```{r}
youtube %>% 
  mutate(number_exclamation = str_count(title, "\\!")) %>% 
  distinct(video_id, .keep_all = TRUE) %>% 
  filter(number_exclamation < 11) %>% 
  ggplot(aes(x = number_exclamation)) +
  geom_histogram(binwidth = 1, 
                 fill = "plum3",
                 color = "black",
                 size = 0.1) +
  xlim(0,10)
  labs(x = "Number of Capital Letters in Title",
       y = "Number of Trending Videos",
       title = "Number of Exclamation Marks (!) in Trending Video Titles",
       caption = "Calculated with distinct video ID's\n Obeserving titles with fewer than 10 exclamations") +
  theme(text = element_text(family = "Times New Roman", color = "grey20")) 
```

## Is there a discernible trend in the usage of question marks?
```{r}
youtube %>% 
  mutate(number_question = str_count(title, "\\?"),
         avg_capitals_per_title_length = str_count(title, "[A-Z]") / str_length(title)) %>% 
  distinct(video_id, .keep_all = TRUE) %>% 
  ggplot(aes(x = number_question)) +
  geom_histogram(binwidth = 1, 
                 fill = "plum3",
                 color = "black",
                 size = 0.1) +
  labs(x = "Number of Question Marks (?) in Title",
       y = "Number of Trending Videos",
       title = "Number of Question Marks (?) in Trending Video Titles",
       caption = "Calculated with distinct video ID's") +
  theme(text = element_text(family = "Times New Roman", color = "grey20")) 
```



## Regional Differences: Exploring Across the Sea

### Number of Trending Videos By Publish Country
```{r}
youtube %>% 
  group_by(publish_country) %>% 
  mutate(total_video=n()) %>% 
  distinct(video_id, .keep_all=TRUE) %>% 
  summarise(distinct_videos=n(),
            total_obs=mean(total_video)) %>% 
  pivot_longer(cols=distinct_videos:total_obs,
               names_to="total_type",
               values_to="number_videos") %>% 
  ggplot(aes(x=total_type, fill=publish_country))+
  scale_fill_discrete(name="Publish Country")+
  geom_bar(aes(weight=number_videos))+
  facet_wrap(~publish_country)+
  labs(title = "Total Number of Videos vs. Total Distinct Videos By Country", 
       y = "Number of Videos", 
       x = "Type of Total")
```

## Are there differences in publishing times throughout the day for trending videos?
```{r youtube_time, include=F}
youtube_time<-youtube %>% 
  mutate(hour=recode(time_frame, "0:00 to 0:59"="1","1:00 to 1:59"="2","2:00 to 2:59"="3",
                                          "3:00 to 3:59"="4","4:00 to 4:59"="5","5:00 to 5:59"="6",
                                          "6:00 to 6:59"="7","7:00 to 7:59"="8","8:00 to 8:59"="9",
                                          "9:00 to 9:59"="10","10:00 to 10:59"="11","11:00 to 11:59"="12",
                                          "12:00 to 12:59"="13","13:00 to 13:59"="14","14:00 to 14:59"="15",
                                          "15:00 to 15:59"="16","16:00 to 16:59"="17","17:00 to 17:59"="18",
                                          "18:00 to 18:59"="19","19:00 to 19:59"="20","20:00 to 20:59"="21",
                                          "21:00 to 21:59"="22","22:00 to 22:59"="23","23:00 to 23:59"="24")) %>% 
  group_by(publish_country) %>% 
  distinct(video_id, .keep_all=TRUE) %>% 
  mutate(distinct_videos=n()) %>% 
  group_by(publish_country, hour) %>% 
  summarise(time_frame_total=n(),
            distinct_videos=mean(distinct_videos)) %>% 
  mutate(time_frame_prop=case_when(publish_country=="CANADA"~time_frame_total/distinct_videos,
                                      publish_country=="FRANCE"~time_frame_total/distinct_videos,
                                      publish_country=="GB"~time_frame_total/distinct_videos,
                                      publish_country=="US"~time_frame_total/distinct_videos))
```

```{r Figure ???}
# Distribution of Publish Time by Country
ggplot(youtube_time, aes(x=hour, fill=publish_country))+
  scale_fill_discrete(name="Publish Country")+
  geom_bar(aes(weight=time_frame_prop))+
  facet_grid(publish_country~.)+
 labs(x="Hour", y="Proportion of Total Distinct Videos Published By Country", title="Distribution of Times When Trending Videos are Published")
```

## Publishing Time throughout week by country?
```{r}
# By Week by Country
youtube %>% 
  group_by(publish_country) %>% 
  distinct(video_id, .keep_all=TRUE) %>% 
  mutate(distinct_videos=n()) %>% 
  group_by(publish_country, published_day_of_week) %>% 
  summarise(time_frame_total=n(),
            distinct_videos=mean(distinct_videos)) %>% 
  mutate(time_frame_prop=case_when(publish_country=="CANADA"~time_frame_total/distinct_videos,
                                      publish_country=="FRANCE"~time_frame_total/distinct_videos,
                                      publish_country=="GB"~time_frame_total/distinct_videos,
                                      publish_country=="US"~time_frame_total/distinct_videos)) %>% 
  ggplot(aes(x=published_day_of_week, fill=publish_country))+
  scale_fill_discrete(name="Publish Country")+
  geom_bar(aes(weight=time_frame_prop))+
  facet_grid(publish_country~.)+
  labs(x="Day of Week", y="Proportion of Total Distinct Videos Published By Country", title="Days of Week When Trending Videos Published", subtitle="by Country")+
  coord_flip()+
  theme(text = element_text(family = "Times New Roman", color = "grey20"))
```

## Length Between Publish Date and Trending Date
```{r}
# Code takes a long time to run?
youtube_trend<-youtube %>%
  select(video_id, publish_date, trending_date) %>%
  group_by(video_id, publish_date) %>%
  summarise(first_trend=min(trending_date)) %>%
  mutate(days_to_trend=(publish_date %--% first_trend)/ddays(1)) %>%
  ungroup()

youtube_trend %>%
  ggplot(aes(x=publish_date, y=days_to_trend, color=days_to_trend))+
  scale_color_continuous(name ="Days to Trending")+
  geom_point(aes(alpha=0.4))+
  labs(title="Days from Published to Trending", x="Publish Date", y="Days to Trending")
```


```{r}
youtube_trend %>% 
  filter(publish_date>="2017-11-14") %>% 
  ggplot(aes(x=publish_date, y=days_to_trend, color=days_to_trend))+
  scale_color_continuous(name="Days to Trending")+
  geom_point(aes(alpha=0.4))+
  labs(title = "Days from Published to Trending",
       subtitle = "from November 14, 2017 to June 14, 2018", 
       x = "Publish Date", 
       y = "Days to Trending")
```


# What is the weird blank line in the middle????
```{r}
youtube %>% 
  arrange(trending_date) %>% group_by(trending_date) %>% summarise(nbr_trending_videos=n()) %>% filter(trending_date>="2018-04-01", trending_date<="2018-04-30") %>% slice(1:10)
```

# Average Number of trending videos in each country each day (For Intro)
```{r}
youtube %>% group_by(publish_country, trending_date) %>% summarise(n=n()) %>% group_by(publish_country) %>% summarise(mean_videos=mean(n))
```


## Numbers and Analytics: Exploring the Statistics of Trending Videos

```{r}
youtube %>% 
  distinct(video_id, .keep_all = TRUE) %>% 
  ggplot(aes(x=views, y=likes)) + geom_point(aes(color = publish_country, alpha=0), 
                                             show.legend = FALSE) + 
  facet_wrap(~publish_country, scales = "free_y") + xlab("Views") + ylab("Likes") + 
  ggtitle("Views vs. Likes in Publish Countries") + 
  theme(text = element_text(family = "Times New Roman", color = "grey20")) + 
  scale_y_continuous(labels = scales::comma) + scale_x_continuous(labels = scales::comma) 
```

```{r}
youtube %>% 
  distinct(video_id, .keep_all = TRUE) %>% 
  filter(category_id == "Music") %>% 
  filter(dislikes < 200000) %>% 
  ggplot(aes(x=views, y=dislikes)) + geom_point(aes(color = publish_country, alpha=0),
                                                show.legend = FALSE) + 
  facet_wrap(~publish_country) + xlab("Views") + ylab("Dislikes") + 
  ggtitle("Views vs. Dislikes in Publish Countries for \nCategory ID: Music") +
  theme(text = element_text(family = "Times New Roman", color = "grey20")) + 
  scale_y_continuous(labels = scales::comma) + scale_x_continuous(labels = scales::comma) + 
  geom_smooth(method = "lm", color="grey20", se=FALSE)
```

```{r}
youtube %>% 
  distinct(video_id, .keep_all = TRUE) %>% 
  filter(category_id == "Music") %>% 
  filter(comment_count < 600000) %>% 
  ggplot(aes(x=likes, y=comment_count)) + geom_point(aes(color = publish_country, alpha=0), 
                                                     show.legend = FALSE) +
  facet_wrap(~publish_country) + xlab("Likes") + ylab("Comment Count") + 
  ggtitle("Likes vs. Comment Count in Publish Countries for \nCategory ID: Music") + 
  theme(text = element_text(family = "Times New Roman", color = "grey20")) + 
  scale_y_continuous(labels = scales::comma) + scale_x_continuous(labels = scales::comma) + 
  geom_smooth(method = "lm", color="grey20", se=FALSE)
```


```{r}
youtube %>% 
  group_by(video_id) %>% 
  mutate(trend_diff = trending_date - lag(trending_date, order_by = trending_date)) %>% 
  # select(trend_diff, trending.date, video_id) %>% 
  arrange(desc(trend_diff)) %>% 
  distinct(video_id, .keep_all=TRUE) %>% 
  filter(trend_diff > 1) %>% 
  filter(category_id == "Entertainment" | category_id == "Music" |
           category_id == "People and Blogs" | category_id == "Comedy" |
           category_id == "News and Politics") %>% 
  ggplot(aes(x=reorder(category_id, trend_diff, FUN=median), y=trend_diff)) + 
  geom_boxplot(aes(fill=category_id), show.legend=FALSE) +
  coord_flip() + xlab("Category ID") + ylab("Time Between Trending Dates") + 
  ggtitle("Days Before Videos Return to \nTrending in each Category") + 
  theme(text = element_text(family = "Times New Roman", color = "grey20")) 
```

```{r}
youtube %>% 
  group_by(video_id) %>% 
  mutate(days_trending = n()) %>% 
  distinct(video_id, .keep_all=TRUE) %>% 
  # select(video_id, days_trending) %>% 
  filter(days_trending <= 100) %>% 
  ggplot(aes(x=days_trending)) + geom_bar(aes(y=..prop..,fill=publish_country), position="dodge", show.legend = FALSE) +
  facet_wrap(~publish_country, scales="free") + xlab("Number of Days Trending") + 
  ylab("Proportion of Videos") + 
  ggtitle("Number of Days Videos Are \nTrending in Publish Countries") + 
  theme(text = element_text(family = "Times New Roman", color = "grey20"))
```

```{r, echo=F}
youtube_table <- youtube %>% 
  group_by(publish_country) %>% 
  distinct(video_id, .keep_all = TRUE) %>% 
  mutate(num_videos = n()) %>% 
  group_by(publish_country) %>% 
  filter(category_id == "Entertainment" | category_id == "Music" |
           category_id == "People and Blogs" | category_id == "Comedy" |
           category_id == "News and Politics") %>% 
  summarise(num_music = sum(ifelse(category_id=="Music", 1, 0)), 
            num_entertainment = sum(ifelse(category_id=="Entertainment", 1, 0)), 
            num_peopleblogs = sum(ifelse(category_id=="People and Blogs", 1, 0)),
            num_comedy = sum(ifelse(category_id=="Comedy", 1, 0)),
            num_newspolitics = sum(ifelse(category_id=="News and Politics", 1, 0)),
            mean_views = round(mean(views)),
            mean_likes = round(mean(likes)),
            mean_dislikes = round(mean(dislikes)),
            mean_comments = round(mean(comment_count)))

reactable(youtube_table, bordered = TRUE, striped = TRUE, columns = list(
  publish_country = colDef(name = "Publish Country"),
  num_music = colDef(name = "Number of Music Videos"),
  num_entertainment = colDef(name = "Number of Entertainment Videos"),
  num_peopleblogs = colDef(name = "Number of People & Blogs Videos"),
  num_comedy = colDef(name = "Number of Comedy Videos"),
  num_newspolitics = colDef(name = "Number of News & Politics Videos"), 
  mean_likes = colDef(name = "Mean Number of Likes"),
  mean_dislikes = colDef(name = "Mean Number of Dislikes"),
  mean_views = colDef(name = "Mean Number of Views"),
  mean_comments = colDef(name = "Mean Number of Comments")))

youtube_table_cat <- youtube %>% 
  distinct(video_id, .keep_all = TRUE) %>% 
  filter(category_id == "Entertainment" | category_id == "Music" |
           category_id == "People and Blogs" | category_id == "Comedy" |
           category_id == "News and Politics") %>% 
  group_by(category_id) %>% 
  summarise(num_videos = n(), 
            mean_views = round(mean(views)),
            mean_likes = round(mean(likes)),
            mean_dislikes = round(mean(dislikes)),
            mean_comments = round(mean(comment_count)))

reactable(youtube_table_cat, bordered = TRUE, striped = TRUE, columns = list(
  category_id = colDef("Category"),
  num_videos = colDef("Number of Videos"),
  mean_views = colDef(name = "Mean Number of Views"),
  mean_comments = colDef(name = "Mean Number of Comments"), 
  mean_likes = colDef(name = "Mean Number of Likes"),
  mean_dislikes = colDef(name = "Mean Number of Dislikes")))


youtube_table_week <- youtube %>% 
  distinct(video_id, .keep_all = TRUE) %>% 
  filter(category_id == "Entertainment" | category_id == "Music" |
           category_id == "People and Blogs" | category_id == "Comedy" |
           category_id == "News and Politics") %>% 
  group_by(published_day_of_week) %>% 
  summarise(num_music = sum(ifelse(category_id=="Music", 1, 0)), 
            num_entertainment = sum(ifelse(category_id=="Entertainment", 1, 0)), 
            num_peopleblogs = sum(ifelse(category_id=="People and Blogs", 1, 0)),
            num_comedy = sum(ifelse(category_id=="Comedy", 1, 0)),
            num_newspolitics = sum(ifelse(category_id=="News and Politics", 1, 0)),
            mean_views = round(mean(views)),
            mean_likes = round(mean(likes)),
            mean_dislikes = round(mean(dislikes)),
            mean_comments = round(mean(comment_count)))

reactable(youtube_table_week, bordered = TRUE, striped = TRUE, columns = list(
  published_day_of_week = colDef(name = "Published Day of Week"),
  num_music = colDef(name = "Number of Music Videos"),
  num_entertainment = colDef(name = "Number of Entertainment Videos"),
  num_peopleblogs = colDef(name = "Number of People & Blogs Videos"),
  num_comedy = colDef(name = "Number of Comedy Videos"),
  num_newspolitics = colDef(name = "Number of News & Politics Videos"),
  mean_views = colDef(name = "Mean Number of Views"),
  mean_comments = colDef(name = "Mean Number of Comments"), 
  mean_likes = colDef(name = "Mean Number of Likes"),
  mean_dislikes = colDef(name = "Mean Number of Dislikes")))

```