---
title: "CHEATING THE YOUTUBE SYSTEM???"
subtitle: "Trending Youtube Video Analysis (NOT CLICKBAIT!!!)"
author: "Group 5: Taylor Buske, Kailee Ervin, Derek Thada, and Saiyothin Vongpanya"
date: "5/1/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(knitr)
library(rmarkdown)
library(ggplot2)
library(forcats)
library(readr)
library(dplyr)
library(tidyr)
library(lubridate)
library(stringr)
library(extrafont)
library(maps)
library(reactable)

my_theme <- function (base_size = 12, base_family = "Times New Roman") {
  theme_gray() +
    theme(
      rect = element_rect(fill = NA, linetype = 0, colour = NA, size = base_size/22),
      text = element_text(family = base_family, face = "plain", color = "black", size = base_size, hjust = 0.5, vjust = 0.5, angle = 0, lineheight = 0.9, margin = margin(), debug = FALSE),
      panel.border = element_blank(),
      panel.background = element_blank(),
      plot.caption = element_text(color = "grey50"),
      legend.position = "right",
      panel.grid.minor.y = element_line(colour = "grey80"), 
      panel.grid.major.x = element_line(colour = "grey80", size = rel(0.25)),
      panel.grid.minor = element_line(size = rel(0.25), colour = "grey80", linetype = "dotted"),
      panel.grid.major.y = element_line(colour = "grey80", size = rel(0.3)),
      plot.title = element_text(face = "bold", size = base_size * 1.5)
    )
}
theme_set(my_theme())
my_pal <- c("#16A08CFF", "plum3", "gold", "#EE0011FF", "#149BEDFF", "lightpink", "lightsalmon", "#A1C720FF", "#FA6B09FF", "paleturquoise", "seagreen", "palevioletred4", "lightpink", "navajowhite2", "lightblue3", "deepskyblue", "coral", "aquamarine")
scale_colour_discrete <- function(...) {scale_colour_manual(..., values = my_pal)}
scale_fill_discrete <- function(...) {scale_fill_manual(..., values = my_pal)}
```

# Introduction
## About our data:
```{r, inlcude=FALSE}
youtube <- read_csv("https://raw.githubusercontent.com/frost2351/stat480/master/youtube.csv")
```

We used trending YouTube data found on Data World that had all trending videos from November 14, 2017, to June 14, 2018, in the four countries of: Canada, France, Great Britain and the United States.

Our data had 161,470 records and the 17 variables: `video_id`, `trending_date`, `title`, `channel_title`, `category_id`, `publish_date`, `time_frame`, `published_day_of_week`, `publish_country`, `tags`, `views`, `likes`, `dislikes`, `comment_count`, `comments_disabled`, `ratings_disabled`, and `video_error_or_removed`.

For our analysis we initially did some data cleaning, inlcuding changing some variable data types (`trending_date` from factor to date type variable), reordering certain factors in more meaningful orders (`published_day_of_week` to start on Monday and end on Sunday), and recoded `category_id` to be the actual category titles. 
```{r, echo= FALSE}
youtube <- youtube %>% 
  mutate(video_id = as.character(video_id),
         trending_date = ydm(as.character(trending_date)),
         title = as.character(title),
         channel_title = as.character(channel_title),
         category_id = recode(as.factor(category_id), "1"="Film & Animation", "2"="Autos & Vehicles",
                              "10"="Music", "15"="Pets and Animals", "17"="Sports", 
                              "19"="Travel &Events", "20"="Gaming", "22"="People and Blogs", 
                              "23"="Comedy", "24"="Entertainment", 
                              "25"="News and Politics", "26"="Howto & Style", "27"="Education",
                              "28"="Science & Technology", "29"="Nonprofits & Activism", "30"="Movies",
                              "43"="Shows", "44"="Trailers"),
         publish_date = dmy(as.character(publish_date)),
         time_frame=factor(time_frame, levels=c("0:00 to 0:59","1:00 to 1:59","2:00 to 2:59","3:00 to 3:59",
                                                "4:00 to 4:59","5:00 to 5:59","6:00 to 6:59","7:00 to 7:59",
                                                "8:00 to 8:59","9:00 to 9:59","10:00 to 10:59","11:00 to 11:59",
                                                "12:00 to 12:59","13:00 to 13:59","14:00 to 14:59",
                                                "15:00 to 15:59","16:00 to 16:59","17:00 to 17:59",
                                                "18:00 to 18:59","19:00 to 19:59","20:00 to 20:59",
                                                "21:00 to 21:59","22:00 to 22:59","23:00 to 23:59")),
         time_frame1=fct_collapse(youtube$time_frame, morning = c("4:00 to 4:59","5:00 to 5:59",
                                                                    "6:00 to 6:59","7:00 to 7:59",
                                                                    "8:00 to 8:59","9:00 to 9:59",
                                                                    "10:00 to 10:59","11:00 to 11:59"), 
                                    afternoon = c("12:00 to 12:59","13:00 to 13:59","14:00 to 14:59",
                                                  "15:00 to 15:59","16:00 to 16:59", "17:00 to 17:59",
                                                  "18:00 to 18:59","19:00 to 19:59"), 
                                    night = c("20:00 to 20:59","21:00 to 21:59","22:00 to 22:59",
                                              "23:00 to 23:59", "0:00 to 0:59","1:00 to 1:59",
                                              "2:00 to 2:59","3:00 to 3:59")),
         published_day_of_week=factor(published_day_of_week, levels=c("Monday", "Tuesday", "Wednesday",
                                                                      "Thursday", "Friday", "Saturday",
                                                                      "Sunday")),
         published_day_of_week1=fct_collapse(youtube$published_day_of_week, 
                                               weekday = c("Monday", "Tuesday", "Wednesday", "Thursday"),
                                               weekend = c("Friday", "Saturday", "Sunday")),
         tags = as.character(tags))


```

## About YouTube's Trending Video List
According to YouTube’s own help page, there are various factors on how a video becomes trending, including:

* Videos must not be misleading, clickbait, or sensational. 
* Most relevant of users and reflective of broad content on platform 
* Shows diversity of creators 

However, these don’t explicitly explain how YouTube’s algorithm works, so we decided to investigate:

**What are the characteristics that make a trending YouTube video?**

In order to investigate this topic, we looked into four areas:
* Dates and Times
* Video Titles
* Video Categories
* Duration and Metrics

# Dates and Times
## About `trending_date`

Each row in our data corresponds to a video on a specific trending date. Therefore, if a video appears multiple times in our dataset, it has been trending for multiple days. 

```{r, echo= FALSE}
youtube %>% 
  group_by(publish_country) %>% 
  mutate(total_video=n()) %>% 
  distinct(video_id, .keep_all=TRUE) %>% 
  summarise(`Distinct Videos`=n(),
            `Total Obs`=mean(total_video)) %>% 
  pivot_longer(cols=`Distinct Videos`:`Total Obs`,
               names_to="total_type",
               values_to="number_videos") %>% 
  ggplot(aes(x=total_type, fill=publish_country))+
  scale_fill_discrete(name="Publish Country")+
  geom_bar(aes(weight=number_videos), show.legend = FALSE)+
  facet_wrap(~publish_country)+
  labs(title="Total Number of Videos \nvs. \nTotal Distinct Videos By Country", y="Number of Videos", x="Type of Total", caption= "Figure 1: Distinct Videos by Country")
```

In Figure 1, I compare the total number of observations of trending videos in each country to their number of distinct videos. As you can see, the total observations for each country are relatively equal but, number of distinct videos is drastically different. For Canada and France, there are more distinct videos compared to Great Britain and the United States. Therefore, while Canada and France have more distinct videos, their videos are not on the trending list for very long. In comparison, Great Britain and the US have less distinct videos that are on the trending list a lot longer. For comparison purposes, I will be using proportions of total distinct videos in each country.

## What are the popular hours to publish?

```{r, echo= FALSE}
youtube %>% 
  mutate(hour=recode(time_frame, "0:00 to 0:59"="1","1:00 to 1:59"="2","2:00 to 2:59"="3",
                                          "3:00 to 3:59"="4","4:00 to 4:59"="5","5:00 to 5:59"="6",
                                          "6:00 to 6:59"="7","7:00 to 7:59"="8","8:00 to 8:59"="9",
                                          "9:00 to 9:59"="10","10:00 to 10:59"="11","11:00 to 11:59"="12",
                                          "12:00 to 12:59"="13","13:00 to 13:59"="14","14:00 to 14:59"="15",
                                          "15:00 to 15:59"="16","16:00 to 16:59"="17","17:00 to 17:59"="18",
                                          "18:00 to 18:59"="19","19:00 to 19:59"="20","20:00 to 20:59"="21",
                                          "21:00 to 21:59"="22","22:00 to 22:59"="23","23:00 to 23:59"="24"),
         hour=as.character(hour),
         hour=as.integer(hour)) %>%
  ggplot(aes(x=publish_country, y=hour))+ 
  geom_boxplot(aes(fill=publish_country), show.legend=FALSE)+
  scale_y_continuous(breaks=seq(0,24,2))+
  scale_x_discrete(limits=c("US", "GB", "FRANCE", "CANADA"))+
  coord_flip()+
  labs(title="Trending Videos Publishing Hour \nby Country", x="Publish Country", y="Hour", caption="Figure 2: Popular Publishing Hours")
```

The variable time_frame is the hour that a video is published. This variable takes into account time zones, so the corresponding time_frame is based on the local time of wherever the video was published.

For easier visualization purposes, We created a new variable called hour that refers to the hour of day the video was published, starting at hour 0 that refers to the time between 12 and 12:59 am.

You can see in Figure 2 that there are no videos published in Hour 0 for any of the countries. Most YouTube creators have set times and days to publish their videos, so it is interesting that none of them release videos exactly at midnight. Looking at the distributions of the boxplots, all countries seem to prefer publishing later in the day as the boxplots are all slightly left skewed. 

Canada and France have the same median publishing hour at around 5pm. However, France is the only country with outliers at hours 1, 2, and 3. Great Britain and the United States also seem to be similar as they have the same median publishing hour at around 4pm. 

## What are the popular days of the week to publish?

```{r, echo= FALSE}
youtube %>% 
  group_by(publish_country) %>% 
  distinct(video_id, .keep_all=TRUE) %>% 
  mutate(distinct_videos=n()) %>% 
  group_by(publish_country, published_day_of_week) %>% 
  summarise(time_frame_total=n(),
            distinct_videos=mean(distinct_videos)) %>% 
  mutate(time_frame_prop=case_when(publish_country=="CANADA"~time_frame_total/distinct_videos,
                                      publish_country=="FRANCE"~time_frame_total/distinct_videos,
                                      publish_country=="GB"~time_frame_total/distinct_videos,
                                      publish_country=="US"~time_frame_total/distinct_videos)) %>% 
  ggplot(aes(x=published_day_of_week, fill=publish_country))+
  scale_fill_discrete(name="Publish Country")+
  geom_bar(aes(weight=time_frame_prop))+
  facet_grid(publish_country~.)+
  labs(x="Day of Week", y="Proportion of Total Distinct Videos Published By Country", title="Days of Week When Trending Videos Published", subtitle="by Country", caption="Figure 3: Popular Days of the Week to Publish")+
  coord_flip()
```

Looking at Figure 3, Canada and France don’t seem to have any significant preferences on which day of the week they publish their trending videos. 

However, both Great Britain and the United States seem to prefer publishing more on weekdays and less on weekends. This result could correspond to the preferences of posting later in the day. As people start getting off work or done with school activities around 4 to 5pm YouTube becomes a popular pastime and creators might take advantage of this time in order to increase their rate of views.

## How long does it take to for a video to become trending?
```{r, echo= FALSE}
youtube_trend<-youtube %>% 
  select(video_id, publish_date, trending_date) %>% 
  group_by(video_id, publish_date) %>% 
  summarise(first_trend=min(trending_date)) %>% 
  mutate(days_to_trend=(publish_date %--% first_trend)/ddays(1)) %>% 
  ungroup()
```

```{r, echo= FALSE}
youtube_trend %>% 
  ggplot(aes(x=publish_date, y=days_to_trend, color=days_to_trend))+
  scale_color_continuous(name ="Days to Trending")+
  geom_point(aes(alpha=0.4))+
  labs(title="Days from Published to Trending", x="Publish Date", y="Days to Trending", caption="Figure 4: How Long from Published to Trending")
```

In Figure 4, We looked at how many days it takes for a video to first appear on the trending list from when it was published.

As a reminder, our data includes all trending videos between November 2017 through June 2018. While our trending date period is relatively narrow, the publishing dates are from 2006 to 2018. Therefore, some videos take a long time to become trending. 

```{r, echo= FALSE}
youtube_trend %>% summarise(mean_days_to_trend=mean(days_to_trend))
```

However, on average it takes a week for a video to become trending once published.

We will take a closer look at publishing dates before and during our trending period that starts on November 14, 2017.

### Taking a closer look: July 23, 2006 to November 14, 2017 
```{r, echo= FALSE}
youtube_old<-youtube_trend %>% 
  filter(publish_date<"2017-11-14")
```

```{r, echo= FALSE}
youtube_old %>% 
  ggplot(aes(x=publish_date, y=days_to_trend, color=days_to_trend))+
  scale_color_continuous(name ="Days to Trending")+
  geom_smooth(method = "lm", color="grey20", alpha=0.3, size=0.7, se=FALSE)+
  geom_point(aes(alpha=0.4))+
  labs(title="Days from Published to Trending", subtitle="July 23, 2006 to November 14, 2017",x="Publish Date", y="Days to Trending", caption="Figure 5: Taking a Closer Look Before Trend Period")
```

When we look at videos that were published before our trending period in Figure 5, we can see a very strong negative linear relationship between when the video was published and how many days it takes for that video to become trending. 

```{r, echo= FALSE}
# Videos before Trend
youtube_trend %>% 
  filter(publish_date<"2017-11-14") %>% 
  summarise(videos_before_trend=n())

# Total Videos
youtube_trend %>% 
  summarise(total_videos=n())
```

However, the number of videos published before November 14, 2017 is only 2% of all the videos published in this dataset so it is a relatively small subset.

### Taking a closer look: November 14, 2017 to June 14, 2018

```{r, echo= FALSE}
youtube_trend %>% 
  filter(publish_date>="2017-11-14") %>% 
  ggplot(aes(x=publish_date, y=days_to_trend, color=days_to_trend))+
  scale_color_continuous(name="Days to Trending")+
  geom_point(aes(alpha=0.4))+
  labs(title="Days from Published to Trending",subtitle="from November 14, 2017 to June 14, 2018", x="Publish Date", y="Days to Trending", caption="Figure 6: Taking a Closer Look During Trending Period")
```

Looking at Figure 6, most videos in this dataset only take a couple days to become trending as seen with the thick dark blue along the x axis. 

```{r, echo= FALSE}
youtube_trend %>% 
  filter(publish_date>="2017-11-14") %>% summarise(mean_days_to_trend=mean(days_to_trend))
```


If you calculate the mean number of days to trending just within the trending period, you only get 1.3 days which is less than the 7 days average number of days to trending for all videos. That means those videos published before the trending period are influencing our average.

There is also a distinct blank line in the data around April.

```{r, echo= FALSE}
youtube %>% 
  arrange(trending_date) %>% group_by(trending_date) %>% summarise(nbr_trending_videos=n()) %>% filter(trending_date>="2018-04-01", trending_date<="2018-04-30") %>% slice(1:10)
```

 There are no trending dates from April 8 to April 13, 2018 in this dataset. We’re not sure if this because this dataset is missing data or for another reason such as a shooting at the YouTube headquarters on April 3, 2018. This event might have caused the offices to close temporarily and for the trending list not to have been able to be updated during that time.

# Video Titles
## Do different countries have different title length in their trending videos?

```{r, echo= FALSE}
youtube %>% 
  distinct(video_id, .keep_all = TRUE) %>% 
  mutate(title_length = str_length(title)) %>% 
  ggplot(aes(x = title_length, 
             y = ..ndensity.., 
             fill = publish_country)) +
  geom_histogram(binwidth=10,
                 show.legend = F) +
  facet_wrap(~publish_country) +
  labs(x = "Length of Title", 
       y = "Proportion of Trending Videos", 
       title = "Length of Titles in Trending Youtube Videos \nin Different Countries",
       caption = "Figure 7: Character Length \nCalculated with distinct video ID's") 
```

Looking at figure 7, we can pair up the countries that have similar distributions. Canada and France are both right skewed, having a tail towards longer videos, while the Great Britain and the United States are fairly normally distributed. This is important to keep in mind for future analysis when analyzing aspects of the title that rely on characters.


## Do different countries have different word length in their titles?

While character length is a useful metric , another important aspect to analyze in their titles is their word length.

```{r, echo= FALSE}
youtube %>% 
  distinct(video_id, .keep_all = TRUE) %>% 
  mutate(words = str_extract_all(title, "[A-z]{1,}"),
         wordcount = lengths(words)) %>% 
  ggplot(aes(x = wordcount, 
             y = ..ndensity.., 
             fill = publish_country)) +
  geom_histogram(binwidth=1,
                 show.legend = F) +
  facet_wrap(~publish_country) +
  labs(x = "Number of Words", 
       y = "Proportion of Trending Videos", 
       title = "Number of Words in Titles of Trending Youtube Videos\n in Different Countries",
       caption = "Figure 8: Word Length \nCalculated with distinct video ID's")
```

Similar to character length, we can pair up the countries in a similar fashion. The most fascinating aspect of the graphs is the large spikes in the 0 word length in both Canada and France. By filtering for these videos I was able to dive deeper in and see what was causing the seemingly random peak. Turns out a majority of these videos are Russian videos and our data set was not compatible to convert Russian characters. This leads to two discoveries:

* A proportion of Canada's and France's trending videos are Russian videos,

* and this large peak is more than likely responsible for the tails in Canada's and France's character length.

## Do different countries use differing amount of capital letters in their trending videos?

While analyzing this metric, we had to consider the varying character lengths for each country. Instead of just using the shear number of capital letters, we use the proportion of capital letters in the title. We believe this better represents the data we are trying to observe. 

```{r, echo= FALSE}
youtube %>% 
  distinct(video_id, .keep_all = TRUE) %>% 
  mutate(avg_capitals_per_title_length = str_count(title, "[A-Z]") / str_length(title)) %>% 
  ggplot(aes(x = avg_capitals_per_title_length, 
             y = ..ndensity.., 
             fill = publish_country)) +
  geom_histogram(binwidth=0.03,
                 show.legend = F) +
  facet_wrap(~publish_country) +
  labs(x = "Proportion of Capital Letters in Title", 
       y = "Proportion of Trending Videos", 
       title = "Proportion of Capital Letters in Titles of Trending \nYoutube Videos in Different Countries",
       caption = "Figure 9: Proportion of Capital Letters \nCalculated with distinct video ID's")
```

Anaylzing the above figure, we can observe that most of the countries have a similar distribution, with the exception of France. While the other countries are similar distributed with a peak around a 20% capital letter proportion, France has a wider peak covering 0-18%. Additionally France is bimodial, repeaking around the 75% proportion mark. Further research lead to no concrete conclusion to why this might occur. 

## Do special characters (! and ?) appear more often in trending videos?
```{r, echo= FALSE}
youtube %>% 
  mutate(number_exclamation = str_count(title, "\\!")) %>% 
  distinct(video_id, .keep_all = TRUE) %>% 
  filter(number_exclamation < 11) %>% 
  ggplot(aes(x = number_exclamation)) +
  geom_histogram(binwidth = 1, 
                 fill = "plum3",
                 color = "black",
                 size = 0.1) +
  xlim(0,10)
  labs(x = "Number of Capital Letters in Title",
       y = "Number of Trending Videos",
       title = "Number of Exclamation Marks (!) in Trending Video Titles",
       caption = "Figure 10: Calculated with distinct video ID's\n Obeserving titles with fewer than 10 exclamations") +
  theme(text = element_text(family = "Times New Roman", color = "grey20")) 
  
youtube %>% 
  mutate(number_question = str_count(title, "\\?"),
         avg_capitals_per_title_length = str_count(title, "[A-Z]") / str_length(title)) %>% 
  distinct(video_id, .keep_all = TRUE) %>% 
  ggplot(aes(x = number_question)) +
  geom_histogram(binwidth = 1, 
                 fill = "plum3",
                 color = "black",
                 size = 0.1) +
  labs(x = "Number of Question Marks (?) in Title",
       y = "Number of Trending Videos",
       title = "Number of Question Marks (?) in Trending Video Titles",
       caption = "Figure 11: Calculated with distinct video ID's") +
  theme(text = element_text(family = "Times New Roman", color = "grey20")) 
```

Looking at the above Figures 10 and 11 we are quickly able to observe that there is not a significant usage of either character in a title. This quick analysis helps us gather that a trending video does not usually associate itself with clickbaity aspects: such as large amounts of capital letters and special characters such as exclamation marks and question marks.

# Video Categories

Our data consists of 18 different video categories. These categories are broad and range from “Pets and Animals” to “Music” to “News and Politics”. It is important to note that each video can only belong to one category, which is chosen by the uploader, not YouTube. Figure 12 below shows the number of instances each category appeared in the daily trending page over our data’s time period. 

```{r, echo= FALSE}
youtube %>% 
  group_by(category_id) %>% 
  mutate(n = n()) %>% 
  ggplot(aes(x = fct_reorder(category_id, n),
             fill = category_id)) + 
  geom_bar(show.legend = FALSE) +
  labs(x = "Category", 
       y = "Instances on Trending Page", 
       title = "Instances on Trending Page in each Category",
       caption="Figure 12") +
  coord_flip() 
```

Thus, we can see that “Entertainment” videos are the most frequent type of videos that appear on the trending list. 

We can also see how the number of instances a category has appeared on the trending list has changed over time as well. Our data consists of trending videos from November of 2017 through June of 2018. Figure 13 below shows the number of trending videos over time in the categories “Entertainment,” “Music,” “People and Blogs,” “Comedy,” and “News and Politics”. These categories are the top five categories from Figure 12, whiched showed the categories with the most instances on the daily trending list.

```{r, echo= FALSE}
left_right <- youtube %>% 
  filter(category_id %in% c("Entertainment", "Music",
                            "People and Blogs", "Comedy",
                            "News and Politics")) %>%
  group_by(category_id, trending_date) %>%
  tally() %>% 
  group_by(category_id) %>% 
  slice(which.min(trending_date), which.max(trending_date))

youtube %>% 
  filter(category_id %in% c("Entertainment", "Music",
                            "People and Blogs", "Comedy",
                            "News and Politics")) %>%
  group_by(category_id, trending_date) %>%
  tally() %>% 
  ggplot() +
  geom_line(aes(x = trending_date,
                y = n,
                color = fct_reorder(category_id, desc(n)))) +
  geom_point(data = left_right,
             aes(x = trending_date,
                 y = n,
                 color = category_id)) +
  geom_line(data = left_right,
            aes(x = trending_date,
                y = n,
                color = category_id),
            linetype = "dashed",
            size = 0.40) +
  labs(x = "Date", 
       y = "Instances on Trending Page", 
       title = "Instances on Trending Page in each Category Time Series",
       color = "Category",
       caption = "Figure 13: Observing the top five categories") +
  scale_fill_discrete(name = "Category") 
```

In analyzing Figure 13, we see that the number of instances of “Entertainment” stays relatively constant over time. Also, we see that the time series for "Entertainment" is the highest on the chart. This makes sense since "Entertainment" was by far the most frequent category to appear on the trending list (as seen in Figure 12). Our second most frequent category was "Music". So, we would expect it to be the second highest on the time series chart, which it is. For “Music” videos, we see an increase starting around March and this increase continues into May. I conducted some additional research to explain this increase. According to what I found, many popular artists release new music around the spring because it gives the songs time to become popular before summer when people are outside more and playing music. This time frame also gives songs time to become popular before summer tours commence. For the bottom three categories of “People and Blogs,” "Comedy," and “News and Politics,” we see that they all appear to initially stay constant and then collectively decrease as "Music" increases. This makes sense because there are only so many spots on the trending list. So, as one category increases, others would have to decrease. 

One metric that YouTube uses to determine which videos are trending is the amount of views the video has. Figure 14 below shows the top five categories with the most views on their videos.

```{r, echo= FALSE}
youtube %>%
  group_by(video_id) %>% 
  slice(which.max(trending_date)) %>% 
  group_by(category_id) %>%
  summarise(total_views = sum(views, na.rm = T)) %>%
  arrange(desc(total_views)) %>%
  top_n(5) %>%
  ggplot(aes(x = fct_reorder(category_id, desc(total_views)),
             fill = category_id)) +
  geom_bar(aes(weight=total_views),
           show.legend = FALSE) +
  labs(x = "Categories", 
       y = "Total Views", 
       title = "Top 5 Categories by Views",
       caption = "Figure 14: Considering the views from the last day trending") +
  scale_y_continuous(labels = scales::comma)
```

In analyzing Figure 14, we see that the top 5 categories are the same as in Figure 2 except that "Sports" has replaced "News and Politics." We see that “Music” has by far the highest amount of views, with the next highest category being "Entertainment." In Figure 12, we found that the most frequent type of trending video was "Entertainment" so this result seems unintuitive. We would expect to see "Entertainment" with the most views since this category appears the most on the trending list. However, according to YouTube, new music releases are some of the most viewed videos on the site, even though they are not on the trending list as often as "Entertainment" videos.  In looking at the most viewed videos in YouTube history, I found that 18 out of the top 20 are “Music” videos. Another explanation for why "Music" has more views than "Entertainment" is that we think people are more likely to replay a "Music" video, which would increase its views. It is not uncommon for someone to play a catchy song a few times, but for "Entertainment" videos, once you watch it, you usually do not get much more satisfaction watching it a second time, or at least not right away. So, this is just another possible reason for why the "Music" category has more views, even though there are more instances of "Entertainment" videos. 

We can also analyze which categories prompt the most fan interaction. We can measure this by analyzing the amount of comments the videos in each category received. Figure 15 shows the top five categories by the amount of comments on their videos.

```{r, echo= FALSE}
youtube %>%
  group_by(video_id) %>% 
  slice(which.max(trending_date)) %>% 
  group_by(category_id) %>%
  summarise(total_comments = sum(comment_count)) %>%
  arrange(desc(total_comments)) %>%
  top_n(5) %>%
  ggplot(aes(x = fct_reorder(category_id, desc(total_comments)),
             fill = category_id)) +
  geom_bar(aes(weight = total_comments),
           show.legend = FALSE) +
  labs(x = "Total Comments", 
       y = "Channel", 
       title = "Top 5 Categories by Comments",
       caption = "Figure 15: Considering the comments from the last day trending") +
  scale_y_continuous(labels = scales::comma)
```

We can see that the categories in Figure 15 are the same as in Figure 13, except "Comedy" and "People and Blogs" have switched places. Also in Figure 15, we see that "Entertainment" has now overtaken "Music." This result may seem unintuitive because we found in Figure 14 that "Music" has the most views. So, the more people watching a video, the more likely, you would think, they would be to leave a comment. However, if we consider the replay scenario mentioned in Figure 14's analysis, we believe that a viewer is unlikely to comment each time they replay a video. Also, another explanation for why "Entertainment" has the highest comment count is because these videos may be more reactionary in nature and may even ask viewers to leave a comment, which a song is unlikely to do.

# Duration and Metrics
```{r, include = F}
youtube %>% 
  group_by(video_id) %>% 
  mutate(days_trending = n()) %>% 
  distinct(video_id, .keep_all=TRUE) %>% 
  filter(days_trending <= 100) %>% 
  group_by(publish_country) %>% 
  summarise(median_days_trend = median(days_trending))
```

```{r, echo=FALSE}
youtube %>% 
  group_by(video_id) %>% 
  mutate(days_trending = n()) %>% 
  distinct(video_id, .keep_all=TRUE) %>% 
  filter(days_trending <= 100) %>% 
  ggplot(aes(x=days_trending)) + geom_bar(aes(y=..prop..,fill=publish_country), position="dodge", show.legend = FALSE) +
  facet_wrap(~publish_country, scales="free") + xlab("Number of Days Trending") + 
  ylab("Proportion of Videos") + 
  ggtitle("Number of Days Videos Are \nTrending in Publish Countries") + 
  labs(caption="Figure 16")+
  theme(text = element_text(family = "Times New Roman", color = "grey20"))
```

Figure 16 shows the number of days videos are trending in each of the four publish countries using the proportion of videos. This graph is also filtered for the number of days trending as less than 100 because there were some very large outliers that made the graphs very difficult to read. As we can see, in Canada and France it appears that the videos they published trend for shorter periods of time - usually less than a week. In Great Britain and the US, however, the videos published tend to trend for longer. While it appears that the majority of these videos trend for two weeks or less, Great Britain videos can trend up to about 50 days, and US videos can trend up to about 60 days. In every country other than the US, the largest proportion of videos trend for one day. In the US, however, there is a spike of the greatest proportion at about one week. The median number of days videos are trending in Canada and France is 1. In Great Britain it is 10, which we see from the curve not tapering off as quickly. In the US, videos trend for a median of 7 days.

```{r, echo = FALSE}
#This only says those who trend for 2 weeks or less
  youtube %>% 
  group_by(video_id) %>% 
  mutate(days_trending = n()) %>% 
  distinct(video_id, .keep_all=TRUE) %>% 
  # select(video_id, days_trending) %>% 
  filter(days_trending <= 14) %>% 
    filter(category_id == "Entertainment" | category_id == "Music" |
             category_id == "People and Blogs" | category_id == "Comedy" |
             category_id == "News and Politics") %>% 
  ggplot(aes(x=days_trending)) + geom_bar(aes(y=..prop..,fill=category_id), position="dodge", 
                                          show.legend = FALSE) +
    facet_wrap(~category_id) +
  xlab("Number of Days Trending") + 
  ylab("Proportion of Videos") + 
  ggtitle("Number of Days Videos Are Trending in \nTop 5 Categories") + 
  labs(caption="Figure 17")+
  theme(text = element_text(family = "Times New Roman", color = "grey20"))
```

Since there were very few videos that were on the trending list for more than two weeks, this graph shows videos that trend for 2 weeks or less. Each of the top 5 categories by number of instances has a peak in number of days trending at 1. The largest proportion of videos in each of these categories trend for one day. Each distribution is heavily right skewed. However, 'People and Blogs', 'Entertainment', and 'News and Politics' have similar distributions. Most of the videos trend for one day, few trend for two days, and fewer trend for three days. There are small proportions of videos that trend for more than three days in these categories. 'Music' and 'Comedy' videos also have similar trends, but slightly larger proportions of these videos trend for longer periods of time. They also still have a spike at one day, but the spike is not as high as in the other three categories. 

```{r, include = F}
youtube %>% 
  group_by(video_id) %>% 
  mutate(trend_diff = trending_date - lag(trending_date, order_by = trending_date)) %>% 
  arrange(desc(trend_diff)) %>% 
  distinct(video_id, .keep_all=TRUE) %>% 
  filter(trend_diff > 1) %>% 
  filter(category_id == "Entertainment" | category_id == "Music" |
           category_id == "People and Blogs" | category_id == "Comedy" |
           category_id == "News and Politics") %>% 
  group_by(category_id) %>% 
  summarise(median_diff = median(trend_diff))
```

```{r, echo=FALSE}
youtube %>% 
  group_by(video_id) %>% 
  mutate(trend_diff = trending_date - lag(trending_date, order_by = trending_date)) %>% 
  arrange(desc(trend_diff)) %>% 
  distinct(video_id, .keep_all=TRUE) %>% 
  filter(trend_diff > 1) %>% 
  filter(category_id == "Entertainment" | category_id == "Music" |
           category_id == "People and Blogs" | category_id == "Comedy" |
           category_id == "News and Politics") %>% 
  ggplot(aes(x=reorder(category_id, trend_diff, FUN=median), y=trend_diff)) + 
  geom_boxplot(aes(fill=category_id), show.legend=FALSE) +
  coord_flip() + xlab("Category ID") + ylab("Time Between Trending Dates") + 
  ggtitle("Days Before Videos Return to \nTrending in each Category") + 
  labs(caption = "Figure 18")+
  theme(text = element_text(family = "Times New Roman", color = "grey20")) 
```

Some videos were on the trending list multiple dates that were not consecutive. This plot does not include consecutive days trending. If a video leaves the trending list, it appears to be likely to come back within only a couple of days. However, there are many outliers as shown on the boxplot. Some videos in the 'Music' category even took about a month to return to the trending list. Videos in the category of 'Music' have the largest spread of time between trending dates. Videos in the category of 'News and Politics' have the smallest spread and least time between trending dates. The top 4 categories on this boxplot have a median number of days before returning to trending of 3 days. In the 'News and Politics' category, the median number of days is 2.

## Numerical Relationships

There were many possible relationships to explore, and many were explored. Following are some of the most interesting associations found.

```{r, echo = FALSE, fig.width=8, fig.height=5}
youtube %>% 
  distinct(video_id, .keep_all = TRUE) %>% 
  ggplot(aes(x=views, y=likes)) + geom_point(aes(color = publish_country, alpha=0), 
                                             show.legend = FALSE) + 
  facet_wrap(~publish_country, scales = "free_y") + 
  xlab("Views") + 
  ylab("Likes") + 
  labs(caption="Figure 19")+
  ggtitle("Views vs. Likes in Publish Countries") + 
  scale_y_continuous(labels = scales::comma) + scale_x_continuous(labels = scales::comma) 
```

We would predict that the more views a video has the more likes it has, and we see this through the upward trends. There is quite a bit of fanning out in each of these graphs. In Canada and France, the association between views and likes is positive and linear. In Great Britain and the US, the association is still positive, but the association is less upward sloping than the other countries. The amount of likes does not increase as much as the amount of views is rising compared to Canada and France. We can’t be sure that this would not be the same for the other countries, however, because Great Britain and the US have the most videos with the largest number of views. If the other countries had more videos with more views, we may see a similar trend.

```{r, echo = FALSE, fig.width=8, fig.height=5}
youtube %>% 
  distinct(video_id, .keep_all = TRUE) %>% 
  filter(category_id == "Music") %>% 
  filter(dislikes < 200000) %>% 
  ggplot(aes(x=views, y=dislikes)) + 
  geom_point(aes(color = publish_country, alpha=0),
                                                show.legend = FALSE) +
  facet_wrap(~publish_country) + 
  xlab("Views") + 
  ylab("Dislikes") + 
  labs(caption="Figure 20")+
  ggtitle("Views vs. Dislikes in Publish Countries for \nCategory ID: Music") +
  theme(text = element_text(family = "Times New Roman", color = "grey20")) + 
  scale_y_continuous(labels = scales::comma) + scale_x_continuous(labels = scales::comma) + 
  geom_smooth(method = "lm", color="grey20", se=FALSE, formula = y~x)
```

We see a positive and fairly linear trend in this figure as the one before, but now only looking at the 'Music' category and dislikes rather than likes. It appears that a video that was published in Great Britain is an outlier with nearly 300,000 dislikes, so it was removed to make the graphs more interpretable. Great Britain has the smallest sloping line - their more viewed videos do not tend to be disliked as much compared to the other three countries. Similarly to the association between views and likes, the trend between views and dislikes is positive. So, the more views a video has, the more dislikes it will have in this category.

```{r, echo = FALSE}
youtube %>% 
  distinct(video_id, .keep_all = TRUE) %>% 
  filter(category_id == "Music") %>% 
  filter(comment_count < 600000) %>% 
  ggplot(aes(x=likes, y=comment_count)) + geom_point(aes(color = publish_country, alpha=0), 
                                                     show.legend = FALSE) +
  facet_wrap(~publish_country) + xlab("Likes") + ylab("Comment Count") + 
  labs(caption="Figure 21")+
  ggtitle("Likes vs. Comment Count in Publish Countries for \nCategory ID: Music") + 
  theme(text = element_text(family = "Times New Roman", color = "grey20")) + 
  scale_y_continuous(labels = scales::comma) + scale_x_continuous(labels = scales::comma) + 
  geom_smooth(method = "lm", color="grey20", se=FALSE, formula = y~x)
```

There is a video that was published in the US with nearly 600,000 comments and 4,000,000 views. This video was removed to make the graph more interpretable. The more likes a video has in the category of 'Music', the more comments there are. The trend is similar in each of the four countries. The lines are positively sloped and pretty linear. Higher number of likes is associated with a higher number of comments. There are very few outliers that stray from the highly linear pattern.


#### Tables - Further Findings?

```{r, include = F}
youtube_table <- youtube %>% 
  group_by(publish_country) %>% 
  distinct(video_id, .keep_all = TRUE) %>% 
  mutate(num_videos = n()) %>% 
  group_by(publish_country) %>% 
  filter(category_id == "Entertainment" | category_id == "Music" |
           category_id == "People and Blogs" | category_id == "Comedy" |
           category_id == "News and Politics") %>% 
  summarise(num_music = sum(ifelse(category_id=="Music", 1, 0)), 
            num_entertainment = sum(ifelse(category_id=="Entertainment", 1, 0)), 
            num_peopleblogs = sum(ifelse(category_id=="People and Blogs", 1, 0)),
            num_comedy = sum(ifelse(category_id=="Comedy", 1, 0)),
            num_newspolitics = sum(ifelse(category_id=="News and Politics", 1, 0)),
            mean_views = round(mean(views)),
            mean_likes = round(mean(likes)),
            mean_dislikes = round(mean(dislikes)),
            mean_comments = round(mean(comment_count)))
```

```{r, echo=FALSE}
reactable(youtube_table, bordered = TRUE, striped = TRUE, columns = list(
  publish_country = colDef(name = "Publish Country"),
  num_music = colDef(name = "Number of Music Videos"),
  num_entertainment = colDef(name = "Number of Entertainment Videos"),
  num_peopleblogs = colDef(name = "Number of People & Blogs Videos"),
  num_comedy = colDef(name = "Number of Comedy Videos"),
  num_newspolitics = colDef(name = "Number of News & Politics Videos"), 
  mean_likes = colDef(name = "Mean Number of Likes"),
  mean_dislikes = colDef(name = "Mean Number of Dislikes"),
  mean_views = colDef(name = "Mean Number of Views"),
  mean_comments = colDef(name = "Mean Number of Comments")))
```

Here we have a summary of some important metrics in the data. France has the largest number of videos in four of the top five most common categories by instances. Canada has the largest number of 'Entertainment' videos. There are also metrics for the mean number of likes, dislikes, comments, and views in each of the four countries. Great Britain leads in mean number of views, likes, dislikes, and comments.

```{r, include = F}
youtube_table_cat <- youtube %>% 
  distinct(video_id, .keep_all = TRUE) %>% 
  filter(category_id == "Entertainment" | category_id == "Music" |
           category_id == "People and Blogs" | category_id == "Comedy" |
           category_id == "News and Politics") %>% 
  group_by(category_id) %>% 
  summarise(num_videos = n(), 
            mean_views = round(mean(views)),
            mean_likes = round(mean(likes)),
            mean_dislikes = round(mean(dislikes)),
            mean_comments = round(mean(comment_count)))
```

```{r, echo=FALSE}
reactable(youtube_table_cat, bordered = TRUE, striped = TRUE, columns = list(
  category_id = colDef("Category"),
  num_videos = colDef("Number of Videos"),
  mean_views = colDef(name = "Mean Number of Views"),
  mean_comments = colDef(name = "Mean Number of Comments"), 
  mean_likes = colDef(name = "Mean Number of Likes"),
  mean_dislikes = colDef(name = "Mean Number of Dislikes")))
```

Looking at the top five categories by number of instances on the trending list, here we have the number of distinct videos trending in that category, and mean number of views, likes, dislikes, and comments. We see the number of 'Entertainment' videos is the highest and the number of 'Music' videos is the lowest. However, we see that 'Music' videos have the highest mean number of views, likes, dislikes, and comments. 'Music' videos appear to attract a lot of viewers and viewer interactions.

```{r, include = F}
youtube_table_week <- youtube %>% 
  distinct(video_id, .keep_all = TRUE) %>% 
  filter(category_id == "Entertainment" | category_id == "Music" |
           category_id == "People and Blogs" | category_id == "Comedy" |
           category_id == "News and Politics") %>% 
  group_by(published_day_of_week) %>% 
  summarise(num_music = sum(ifelse(category_id=="Music", 1, 0)), 
            num_entertainment = sum(ifelse(category_id=="Entertainment", 1, 0)), 
            num_peopleblogs = sum(ifelse(category_id=="People and Blogs", 1, 0)),
            num_comedy = sum(ifelse(category_id=="Comedy", 1, 0)),
            num_newspolitics = sum(ifelse(category_id=="News and Politics", 1, 0)),
            mean_views = round(mean(views)),
            mean_likes = round(mean(likes)),
            mean_dislikes = round(mean(dislikes)),
            mean_comments = round(mean(comment_count)))
```

```{r, echo=FALSE}
reactable(youtube_table_week, bordered = TRUE, striped = TRUE, columns = list(
  published_day_of_week = colDef(name = "Published Day of Week"),
  num_music = colDef(name = "Number of Music Videos"),
  num_entertainment = colDef(name = "Number of Entertainment Videos"),
  num_peopleblogs = colDef(name = "Number of People & Blogs Videos"),
  num_comedy = colDef(name = "Number of Comedy Videos"),
  num_newspolitics = colDef(name = "Number of News & Politics Videos"),
  mean_views = colDef(name = "Mean Number of Views"),
  mean_comments = colDef(name = "Mean Number of Comments"), 
  mean_likes = colDef(name = "Mean Number of Likes"),
  mean_dislikes = colDef(name = "Mean Number of Dislikes")))

```

This table shows important metrics for trending videos that were published on each day of the week in the top five categories by number of instances on the trending list. For example, in each of the top 5 categories other than 'News and Politics', the most videos are posted on Fridays. The most trending 'News and Politics' videos are published on Mondays. Also, the largest mean number of views, likes, and comments happen on Fridays, while the largest mean number of dislikes happen on Sundays. 

# Conclusions
* Entertainment and Music videos appear on the trending list the most
  + Implies that these categories are the most popular on the site
* Most videos take about a week to begin trending 
  + A trending video will become viral quickly from its release
  + Implies that a fanbase helps bolster getting the videos seen quicker
* Videos trend for longer in Great Britain and the US 
  + Reinforces the idea that similar channels and creators repopulate the trending page
* A lot of factors may have small influences over a trending video 
  - Publish Country 
  - Category 
  - Date & Time published 

# Future Directions
* Video Tags 
  + One of the few truly "messy" data in our data set
  + A lot of different components need to be parsed out to get workable data
* Having non-trending video data
  + Allows us to do a much more significant analysis on how to make a trending video
  + Would allow for much stronger comparisons
* Analyze more countries
  + YouTube is available all across the globe, not just in the 4 countries we analyze
  + Global data may allow for more inferences to be made on how a video becomes trending
* Evaluate a longer time period 
  + Get a more concrete analysis on the changes that happened over time
* Stronger analysis on titles 
  + No strong conclusions were drawn from the titles
* Parsing words or phrases to observe trends 
  + Need a way to deal with the Russian videos
  + Not enough time to clean the titles to make parsing it usuable


# References
* YouTube Info on Trending List: https://support.google.com/youtube/answer/7239739?hl=en 
* YouTube Headquarters'Shooting: https://www.bbc.com/news/world-us-canada-43645812 
* Information on Music Releases: https://musicconsultant.com/music-career/when-is-the-best-time-of-year-to-release-an-album-or-ep/#.XqiMbUBFzIU and https://www.hypebot.com/hypebot/2019/12/best-and-worst-months-to-release-new-music.html 
* YouTube Video API category_id list: https://gist.github.com/dgp/1b24bf2961521bd75d6c 
* YouTube dataset: https://data.world/iamdilan/youtube-dataset 


